The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-01-22 src/ms.tex
Apply suggestions from code review
Co-authored-by: Lia Corrales <lia.astro@gmail.com>

All affiliated packages are listed on the astropy 
website\footnote{\url{https://www.astropy.org/affiliated}}. Since 
\cite{astropy:2018}, there has been an explosion of gravitational-wave affiliated 
packages including: \texttt{PyCBC} for exploring GW sources, lenstronomy for 
modeling strong gravitational lenses, \texttt{ligo.skymap} for visualizing GW 
probability maps, and \texttt{EinsteinPy} for general relativity and gravitational astronomy. 
There have been several packages related to HEALPix: \texttt{astropy-healpix}
for a BSD licensed HEALPix implementation, and \texttt{mocpy} for Multi-Order 
Coverage maps. \texttt{astroalign} as been introduced for astrometric registration, 
and \texttt{python-cpl} has been added for ESO pipelines and VLT data products. 
For ground-based astronomy, \texttt{baseband} has added IO capabilities for 
VLBI, and \texttt{SpectraPy} brings slit spectroscopy to the astropy ecosystem. 
We have \texttt{agnpy} for AGN astrophysics, and \texttt{statmorph} for fitting galactic 
morphological diagnostics. \texttt{saba} gives an interface to sherpa's fitting 
routines. For modeling interstellar dust extinction we have 
\texttt{dust_extinction}, and for extracting features from time series we 
have \texttt{feets}. We also have \texttt{corral} for managing data intensive parallel 
pipelines. \texttt{BayesicFitting} provides an interface for generic Bayesian 
inference, and \texttt{sbpy} does calculations for asteroid and cometary 
astrophysics. Finally, \texttt{synphot} provides an interface for synthetic photometry.

Several of the affiliated packages described in [reference to paper 2] 
have had substantial improvements. \texttt{Astroquery} [reference] has added access 
to roughly a dozen new missions and data services, including JWST, and the 
project has switched to a continuous release model. Every time a change 
is committed to the main development branch it is published on the Python 
Package Index (PyPI, ref) and available for installation. Formal releases 
are still done a couple of times per year. \texttt{Photutils} released its first 
stable version, indicating that the API will change less frequently, and 
there have been several significant performance improvements.  \texttt{CCDproc} 
also released a new major version, bringing better performance to some 
image combination operations. The \texttt{regions} package, for manipulating 
ds9-style [ref] region definitions, added new ways to manipulate regions 
and introduced new region types. \texttt{Reproject}’s major new feature is a 
function to align and co-add images to create a mosaic; better support 
for parallelization was also added. \texttt{Specutils}, the package that defines 
containers for 1D and 2D spectra, also had its first stable rele and 
the addition of classes to read JWST data.

@@

\begin{itemize}
\item Current status and scope
\item Vision for the future (review what we said in v2.0 paper)
\item User forums and engagement with user base
\item Workshops
\end{itemize}

>>>>

@@

\subsubsection{Current status and scope}

{\it Learn Astropy} is an umbrella term that acknowledges the broad educational efforts made by the Astropy Project, which are led by the Astropy Learn Team. 
As described in \textsl{[Astropy-ii paper]}, Learn Astropy content has been categorized along four axes: documentation, which is contained within the python code base; tutorials, consisting of Jupyter Notebook lessons that are published in HTML format online; guides, which are a series of lessons providing a foundational resource for performing a certain type of astronomical analyses; and examples, which are snippets of code that showcase a short task that can be performed with Astropy or its affiliated package. 
These categories continue to drive both the infrastructure choices and the appearance of the Learn Astropy website. Additionally, the Astropy Learn Team has expanded its efforts
in user engagement and in conducting workshops. The Learn team meets
weekly to drive these issues forward.

\begin{itemize}

\item {\bf Learn Astropy website:} The site has a new search interface powered by
Algolia which is used by major shopping sites. This work has been
supported in part by the Dunlap grant. Tutorials have been added and
a {\it ccdproc} Jupyter book has been completed.

\item {\bf User forums:} The Astropy Project has historically maintained a number of avenues for users to seek help or gain access to the developer community. This includes the Astropy mailing list, the Python in Astronomy Facebook group, and the Astropy Slack workspace. The Moore Foundation grant allows this Slack space to
be on a paid plan; additionally NumFOCUS has negotiated a special rate for
open-source projets. While the Slack and Astropy-dev mailing lists are primarily used to discuss the project direction and updates, it was noted that the use of a Facebook community could present a barrier to open source. The Astropy Project identified a need for a public, archived, searchable, and easily-to-navigate interface for users to ask for help Accordingly, we have commissioned 
a Discourse site which is more open than the Facebook group and more user-oriented
than Astropy Slack. A benefit of the Moore Foundation grant is that Astropy
developers are able to invoice as independent contractors the time they
spend helping users on these forums.

\item {\bf Workshops:} the Project has been conducting workshops at winter meetings
of the American Astronomical Society since AAS 225 in January 2015. Up to the
start of the coronavirus pandemic, these were full-day in-person workshops with
as many as 90 participants and a dozen facilitators from the project. During the
pandemic, these workshops were moved to an online format and split into
basic and advanced sessions. Additionally, beginning with the AAS 238 online
meeting, the workshops have been expanded to Summer AAS meetings. The Learn
team finds that the workshop audience is best found as AAS meetings as opposed
to more general Python meetings.

The Astropy Project provided another mode of community engagement at AAS Meetings 235 and 237 by organizing a NumFOCUS Sponsored Projects booth in the AAS Exhibit Hall. Funding for the exhibit hall was provided alternately by NumFOCUS and later by the Moore Foundation funding (\textbf{check this}). The booth hosted a series of Q\&A special sessions during AAS 235 and webinars during the virtual AAS 237 meetings, to provide the general astronomy community information and access to experts on a variety of open source astronomical tools.
\end{itemize}

\subsubsection{Learn vision for the future}

The Learn plan going forward is to continue to improve the tutorial website;
to index third-party tutorials; to promote user engagement on the Discourse
site; and to look for supportable opportunities to expand the reach of
Astropy workshops beyond the American Astronomical Society meetings.
